---
title: "Homework 4"
author: "Charles Hanks"
date: "02/1/2023"
output: 
  html_document:
    df_print: kable
    fig_width: 11
    fig_height: 8
---

**Directions:**

Please turn in **both** a knitted HTML file *and* your Rmd file on WISE.

Good luck!

# 1. Setup (1pt)

Change the author of this RMD file to be yourself and modify the below code so that you can successfully load the 'wine.rds' data file from your own computer.

```{r setup, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
knitr::opts_knit$set(root.dir = "/Users/charleshanks/Desktop/MSDS/SPRING_23/ML")
library(tidyverse)
library(caret)
library(naivebayes)
library(tidytext)
wine = read_rds("pinot.rds")
names(wine)[names(wine) == 'id'] = 'ID'

```

# 2. Conditional Probability (3pts)

Calculate $P(Burgundy | Fruit)$

...i.e. the probability that a Pinot comes from Burgundy given it has the word 'fruit' in the description.

```{r}
#doc term matrix: 
df <- wine %>%
unnest_tokens(word, description) %>%
anti_join(stop_words) %>% # get rid of stop words
filter(word != "wine") %>%
filter(word != "pinot") %>%
count(ID, word) %>%
group_by(ID) %>%
mutate(freq = n/sum(n)) %>%
mutate(exists = (n>0)) %>%
ungroup %>%
group_by(word) %>%
mutate(total = sum(n))

num_fruit = df %>% filter(word == "fruit") %>% group_by(ID) %>% nrow() 

#3724 wines have fruit in description

total_wine = nrow(wine)
#8380


p_f = num_fruit/total_wine
#.444 probability that there is fruit is description 

#probability that wine is burgundy
p_b = wine %>% filter(province == "Burgundy") %>% count()/total_wine
#.142

#subset with only burgundy wines 
burg_province = wine %>% filter(province == "Burgundy") 

#burgundy wine term matrix
bp_df <- burg_province %>%
unnest_tokens(word, description) %>%
anti_join(stop_words) %>% 
filter(word != "wine") %>%
filter(word != "pinot") %>%
count(ID, word)

fruit_burg = bp_df %>% filter(word == "fruit")
#520 burgundy have fruit in the description 
#So P(Fruit | Burgundy)
p_fg = (nrow(fruit_burg))/(nrow(burg_province))
#.439

# P(Burgundy | Fruit) = P(Fruit | Burgundy) * P(Burgundy) / P(Fruit)
answer = round((p_b*p_fg)/(p_f),2)
answer


```

ANSWER: P(Burgundy | Fruit) = .14




# 3. Naive Bayes Algorithm (4pts)

1. Train a naive bayes algorithm to classify a wine's province,
2. using 80% of your data,
3. three features engineered from the description
4. and 5-fold cross validation.
5. Report Kappa after using your model to predict provinces in the holdout (test) sample.

```{r}
wine2 = wine %>% mutate(lprice = log(price),
                    fyear = as.factor(year),
                    oak = as.integer(str_detect(description, "[Oo]ak")),
                    earth = as.integer(str_detect(description, "[Ee]arth")),
                    cherry = as.integer(str_detect(description, "[Cc]herry")),
                    choc = as.integer(str_detect(description, "[Cc]hocolate")),
                    acidity = as.integer(str_detect(description, "[Aa]cidity")),
                    nose = as.integer(str_detect(description, "[Nn]ose")),
                    palate = as.integer(str_detect(description, "[Pp]alate")),
                    chocolate = as.integer(str_detect(description,"[Cc]hocolate")),
                    tart = as.integer(str_detect(description,"[Tt]art")),
                    brisk = as.integer(str_detect(description,"[Bb]risk")),
                    bramble = as.integer(str_detect(description,"[Bb]ramble")),
                    aging = as.integer(str_detect(description, "[Aa]ging")),
                    savory =as.integer(str_detect(description, "[Ss]avory")),
                    clover = as.integer(str_detect(description, "[Cc]love")),
                    aromas = as.integer(str_detect(description, "[Aa]romas")),
                    fruits = as.integer(str_detect(description, "[Ff]ruits")),
                    nose = as.integer(str_detect(description, "[Nn]ose")),
                    points_greater_95 = points >=95, 
                    points_less_90 = points <= 90,
                    price_greater_4 = lprice >= 4, 
                    price_between_4_3 = lprice < 4 & lprice >= 3, 
                    price_less_3 = lprice < 3,
                    before_2010 = year < 2010, 
                    beween_2010_2015 = (year >=2010 & year <= 2015),
                    between_2015_2020 = (year > 2015 & year <= 2020)) %>%
                    select(-ID,-price,-description)


set.seed(504)

wine_index <- createDataPartition(wine2$province, p = 0.8, list = FALSE)
train <- wine2[ wine_index, ]
test <- wine2[-wine_index, ]

control <- trainControl(method = "cv", number = 5)

fit <- train(province ~ .,             
	data = train,              
	method = "knn",             
	tuneLength = 15,             
	trControl = control)

fit
confusionMatrix(predict(fit,test), factor(test$province)) 

```

ANSWER: kappa = .517


# 4. Frequency differences (2pts)

List the three words that most distinguish New York Pinots from all other Pinots.

```{r}
wtxt <- wine %>% 
  unnest_tokens(word, description) %>% 
  anti_join(stop_words) %>% 
  filter(str_detect(string = word, pattern = "[a-z+]")) %>%  # get rid weird non alphas
  filter(str_length(word)>3) %>%  # get rid of strings shorter than 3 characters
  group_by(word) %>% 
  mutate(total=n()) %>% 
  ungroup()


#getting word frequency for "all other pinots"
not_ny = wtxt %>% filter(province != "New_York") %>%
  filter(!(word %in% c("wine","pinot","drink","noir","vineyard","palate","notes","flavors","bottling","bottle","finish"))) %>% 
  filter(total > 400) %>% 
  group_by(province, word) %>%
  count() %>% 
  group_by(province) %>% 
  mutate(proportion = n / sum(n)) %>% 
  pivot_wider(id_cols = word, names_from = province, values_from = proportion)

#changing NAs to Os
not_ny[is.na(not_ny)] = 0

#getting mean frequency of the 5 provinces 
not_ny$avg_prop = rowMeans(not_ny[,-1])


#getting word frequency for new york pinots
ny = wtxt %>% filter(province == "New_York") %>%
  filter(!(word %in% c("wine","pinot","drink","noir","vineyard","palate","notes","flavors","bottling","bottle","finish"))) %>% 
  filter(total > 400) %>% 
  group_by(province, word) %>%
  count() %>% 
  group_by(province) %>% 
  mutate(proportion = n / sum(n)) %>% 
  pivot_wider(id_cols = word, names_from = province, values_from = proportion)

#checking for any NAs in ny subset 
which(is.na(ny))

#joining ny & not_ny
ny_vs_rest = ny %>% left_join(select(not_ny, word, avg_prop), by = "word") 

#creating word freq difference between ny and rest of pinots, ordering by top difference, grabbing top 3 words.
answer = ny_vs_rest %>% mutate(diff = New_York -avg_prop) %>% arrange(desc(diff)) %>% top_n(3) 

as.list(answer[,1])
```


ANSWER: cherry, tannins, concentrated
