---
title: "Homework 4"
author: ""
date: "02/1/2023"
output: 
  html_document:
    df_print: kable
    fig_width: 11
    fig_height: 8
---

**Directions:**

Please turn in **both** a knitted HTML file *and* your Rmd file on WISE.

Good luck!

# 1. Setup (1pt)

Change the author of this RMD file to be yourself and modify the below code so that you can successfully load the 'wine.rds' data file from your own computer.

```{r setup, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
knitr::opts_knit$set(root.dir = "/Users/charleshanks/Desktop/MSDS/SPRING_23/ML")
library(tidyverse)
library(caret)
library(naivebayes)
wine = read_rds("pinot.rds")
```

# 2. Conditional Probability (3pts)

Calculate $P(Burgundy | Fruit)$

...i.e. the probability that a Pinot comes from Burgundy given it has the word 'fruit' in the description.

```{r}
#Given that it has the word fruit in the description, what is probability that e


wine = wine %>% mutate(fruit = as.integer(str_detect(description, "[Ff]ruit")))

num_fruit = sum(wine$fruit)
#4824 wines have fruit in description
total_wine = nrow(wine)
#8380
fruit_prob = num_fruit/total_wine
#.5756
#probability that wine is burgundy
burg_prob = wine %>% filter(province == "Burgundy") %>% count()/total_wine
#.142

burg_province = wine %>% filter(province == "Burgundy") %>% mutate(fruit = as.integer(str_detect(description,"[Ff]ruit")))
burg_wines_with_fruit = sum(burg_province$fruit)
#1193 burgundy pinots in ds 
#1054 burgundy wines contain word fruit in description. 
#so P(Fruit | Burgundy) = 
1054/1193

prob_burg_given_fruit = (.883*.142)/.576


```

$$ P({\rm Thing 1}~|~{\rm Thing 2}) = \frac{P({\rm Thing 2}~|~ \rm{Thing 1})P(\rm{Thing 1})}{P({\rm Thing 2})} $$
(P(Burgundy | Fruit) = P(Fruit | Burgundy)*P(Burgundy))/P(Thing2)

P(fruit | Burgundy) = .883
P(Burgundy) = .142
P(Fruit) = .576

.883*,142

The probability that a Pinot comes from Burgundy given that it has the word "fruit" in the description is .218. 




# 3. Naive Bayes Algorithm (4pts)

1. Train a naive bayes algorithm to classify a wine's province,
2. using 80% of your data,
3. three features engineered from the description
4. and 5-fold cross validation.
5. Report Kappa after using your model to predict provinces in the holdout (test) sample.

```{r}

ds_nb = wine %>% mutate(acidity = as.integer(str_detect(description, "[Aa]cidity")),
                       nose = as.integer(str_detect(description, "[Nn]ose")),
                       palate = as.integer(str_detect(description, "[Pp]alate"))) %>% 
                select(province, acidity,nose, palate )


wine_index <- createDataPartition(ds_nb$province, p = 0.8, list = FALSE)
train <- ds_nb[ wine_index, ]
test <- ds_nb[-wine_index, ]

control <- trainControl(method = "cv", number = 5)

nbm <- train(province ~ .,
             data = train, 
             method = "naive_bayes",
             tuneGrid = expand.grid(usekernel = c(T,F), laplace = T, adjust = T),
             metric = "Kappa",
             trControl = control)

nbm

confusionMatrix(predict(nbm,test), factor(test$province))

```


```

# 4. Frequency differences (2pts)

List the three words that most distinguish New York Pinots from all other Pinots.

```{r}

#FEATURE ENGINEERING 
df %>% filter(starts_with("herb") )

provinces = c("Burgundy", "California", "Casablanca_Valley","Marlborough", 
                 "New_York", "Oregon")

oregon = wine %>% filter(province == "Oregon")
  
oregon_df = oregon %>%
  unnest_tokens(word, description) %>%
  anti_join(stop_words) %>% # get rid of stop words
  filter(word != "wine") %>%
  filter(word != "pinot") %>%
  count(ID, word) %>% 
  group_by(ID) %>% 
  mutate(freq = n/sum(n)) %>% 
  mutate(exists = (n>0)) %>% 
  ungroup %>% 
  group_by(word) %>% 
  mutate(total = sum(n))

Marlborough_top25words = wine %>% filter(province == "Marlborough") %>%
  unnest_tokens(word, description) %>%
  anti_join(stop_words) %>% # get rid of stop words
  filter(word != "wine") %>%
  filter(word != "pinot") %>%
  count(ID, word) %>% 
  group_by(ID) %>% 
  mutate(freq = n/sum(n)) %>% 
  mutate(exists = (n>0)) %>% 
  ungroup %>% 
  group_by(word) %>% 
  mutate(total = sum(n)) %>% 
  count(word) %>% 
  arrange(desc(n)) %>% 
  head(25)
  
 = oregon_df %>% 
  count(word) %>%
  arrange(desc(n)) %>% 
  head(25) %>% select(word)


colnames(california_top25words) = c("ca_words")
colnames(burgundy_top25words) =   c("burg_words")
colnames(Marlborough_top25words) = c("marl_words")
colnames(casablanca_top25words) = c("casa_words")
colnames(new_york_top25words) = c("ny_words")
colnames(oregon_top25words) = c("or_words")

california_top25words$index = 1:nrow(california_top25words)

l1 = list(burgundy_top25words, Marlborough_top25words,casablanca_top25words,new_york_top25words,oregon_top25words)




top_25_by_province = cbind(california_top25words,burgundy_top25words,Marlborough_top25words,casablanca_top25words,new_york_top25words,oregon_top25words)

view(top_25_by_province)

unique_words1 = unique(unlist(top_25_by_province[,1]))
unique_words2 = unique(unlist(top_25_by_province[,2]))
unique_words3 = unique(unlist(top_25_by_province[,3]))
only_in_col1 = setdiff(unique_words1,unique_words2)
only_in_col2 = setdiff(unique_words2,unique_words1)

list(only_in_col1 = only_in_col1, only_in_col2 = only_in_col2)
?setdiff()

wine %>% filter(str_detect(description, "crisp")) %>% group_by(province) %>% count()

wine %>% group_by(province) %>% count()

burgundy_top25words
brisk, bramble

```

