---
title: "ML4_group_activity"
output: pdf_document
date: "2023-02-01"
---

## Group exercise

Use the top words by province to...

1. Engineer more features that capture the essence of Casablanca, Marlborough and New York
2. Look for difference between California and Oregon
3. Use what you find to run naive Bayes models that achieve a Kappa that approaches 0.5

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/Users/charleshanks/Desktop/MSDS/SPRING_23/ML")
library(tidyverse)
library(caret)
library(naivebayes)
library(fastDummies)

wine = read_rds("pinot.rds")
names(wine)[names(wine) == 'id'] = 'ID'
```



```{r}
wine
```

Document term matrix: 

```{r}
df <- wine %>%
  unnest_tokens(word, description) %>%
  anti_join(stop_words) %>% # get rid of stop words
  filter(word != "wine") %>%
  filter(word != "pinot") %>%
  count(ID, word) %>% 
  group_by(ID) %>% 
  mutate(freq = n/sum(n)) %>% 
  mutate(exists = (n>0)) %>% 
  ungroup %>% 
  group_by(word) %>% 
  mutate(total = sum(n))
```

Top words in database:

```{r}
df %>% 
  count(word) %>%
  arrange(desc(n)) %>% 
  head(25)
```




Combining words with over 1000 occurences in dataset with ID and province from wine 
```{r}
wino <- df %>% 
  filter(total > 800) %>% 
  pivot_wider(id_cols = ID, names_from = word, values_from = exists, values_fill = list(exists=0)) %>% 
  merge(select(wine,ID, province), all.y=TRUE) #%>% 
  #drop_na()

#wino <- merge(select(wine,ID, province), wino, by="ID", all.x=TRUE) %>%
#  arrange(ID)
#View(wino)
wino <- replace(wino, is.na(wino), FALSE)
```
run model
```{r}
set.seed(504)
wine_index <- createDataPartition(wino$province, p = 0.80, list = FALSE)
train <- wino[ wine_index, ]
test <- wino[-wine_index, ]

fit <- train(province ~ .,
             data = train, 
             method = "naive_bayes",
             tuneGrid = expand.grid(usekernel = c(T,F), laplace = T, adjust = T),
             metric = "Kappa",
             trControl = trainControl(method = "cv"))

confusionMatrix(predict(fit, test),factor(test$province))

#kappa = .39
```

## Maybe we can find words associated with our sparse provinces?

```{r}
df %>% 
  left_join(select(wine, ID, province), by = "ID") %>% 
  count(province, word) %>%
  group_by(province) %>% 
  top_n(5,n) %>% 
  arrange(province, desc(n))

```


#adding interaction terms for our sparse provinces 
```{r}
#adding interaction terms
wino = wino %>% mutate(cv_flavors_aromas = flavors*aromas, 
                       m_drink_cherry = drink*cherry, 
                       ny_cherry_noir = cherry*noir)

#adding fruit*chocolate for oregon:
choc = wine %>% mutate(
    choc = str_detect(description, "chocolate")) %>% 
    select(ID,choc)

#taking out interactions terms: 
wino = wino %>% select(-cv_flavors_aromas,-m_drink_cherry,-ny_cherry_noir)

#adding new feature from wine ds: 
noir = wine %>% mutate(
    noir = str_detect(description, "noir")) %>% 
    select(ID,noir)

#merging new feature with wino 
wino = wino %>% 
  left_join(noir, by = "ID")

concentrated = wine %>% mutate(
    concentrated = str_detect(description, "concentrated")) %>% 
    select(ID,concentrated)

#merging new feature with wino 
wino = wino %>% 
  left_join(choc, by = "ID")


```

adding texture, aging
```{r}
aging_texture = wine %>% mutate(
    aging = str_detect(description, "aging"),
    texture = str_detect(description, "texture")) %>% 
    select(ID,aging,texture)

#merging new feature with wino 
wino = wino %>% 
  left_join(aging_texture, by = "ID")
```


run model with interaction terms : 
```{r}
wine_index <- createDataPartition(wino$province, p = 0.80, list = FALSE)
train <- wino[ wine_index, ]
test <- wino[-wine_index, ]

fit <- train(province ~ .,
             data = train, 
             method = "naive_bayes",
             tuneGrid = expand.grid(usekernel = c(T,F), laplace = T, adjust = T),
             metric = "Kappa",
             trControl = trainControl(method = "cv"))

confusionMatrix(predict(fit, test),factor(test$province))
```

differentiating ny from california

```{r}
wtxt %>% 
    filter(province=="New_York" | province=="California") %>% 
    filter(!(word %in% c("wine","pinot","drink","noir","vineyard","palate","notes","flavors","bottling"))) %>% 
    filter(total > 400) %>% 
    group_by(province, word) %>%
    count() %>% 
    group_by(province) %>% 
    mutate(proportion = n / sum(n)) %>% 
    pivot_wider(id_cols = word, names_from = province, values_from = proportion) %>% 
    ggplot(aes(x = New_York, y = California, color = abs(New_York - California))) +
    geom_abline(color = "gray40", lty = 2) +
    geom_jitter(alpha = 0.1, size = 2.5, width = 0.3, height = 0.3) +
    geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
    scale_x_log10(labels = percent_format()) +
    scale_y_log10(labels = percent_format()) +
    scale_color_gradient(limits = c(0, 0.001), low = "darkslategray4", high = "gray75") +
    theme(legend.position="none") +
    labs(x = "New York", y = "California", title = "Words describing Pinot Noir from NY and CA")
```
new york from casablanca
```{r}
wtxt %>% 
    filter(province=="New_York" | province=="Casablanca_Valley") %>% 
    filter(!(word %in% c("wine","pinot","drink","noir","vineyard","palate","notes","flavors","bottling"))) %>% 
    filter(total > 400) %>% 
    group_by(province, word) %>%
    count() %>% 
    group_by(province) %>% 
    mutate(proportion = n / sum(n)) %>% 
    pivot_wider(id_cols = word, names_from = province, values_from = proportion) %>% 
    ggplot(aes(x = New_York, y = Casablanca_Valley, color = abs(New_York - Casablanca_Valley))) +
    geom_abline(color = "gray40", lty = 2) +
    geom_jitter(alpha = 0.1, size = 2.5, width = 0.3, height = 0.3) +
    geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
    scale_x_log10(labels = percent_format()) +
    scale_y_log10(labels = percent_format()) +
    scale_color_gradient(limits = c(0, 0.001), low = "darkslategray4", high = "gray75") +
    theme(legend.position="none") +
    labs(x = "New York", y = "Casablanca_Valley", title = "Words describing Pinot Noir from NY and Casablanca_Valley")
```
```{r}


```


